Modeling Galaxy Spectral Energy Distributions in the SDSS Spectroscopic Survey

### 2002-02-11 docs/paper/paper.tex
added
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@33 75dff457-8751-0410-ab1a-cd701654ce6b

We implemented this system and are publishing it in order to prepare
the reader for upcoming papers which will rely heavily on the
reliability of the SED fits described here.  However, having
implemented the fits, there is no reason not to share them with the
community; thus, this paper serves the secondary purpose of describing
a piece of software being released to the astronomical community. The
computer software created is a C shared object library, around which
we have written both stand-alone C programs and IDL routines. We have
made the source code available publicly, through a URL on the World
Wide Web and through a public CVS repository. Improvements or ports to
other languages implemented by users may be incorporated into the code
upon request. The conditions of use for the code are that this paper
is cited in any resulting refereed journal article and that the
version of the code used is specified in any such paper. The version
of the code used to make the figures for this paper is {\tt kcorrect
v1_2}.

@@

Our task is to recover a model for the galaxy SED from broad-band
photometric measurements. Since a set of broad band galaxy fluxes does
not correspond uniquely to a particular SED, and because galaxy SEDs
are known to have significant structure over wavelength ranges small
compared to our bandpass, this task is an ill-posed, inverse
problem. However, we are not completely ignorant about the forms which
galaxy SEDs take, so we can attempt to use what we know about galaxy
SEDs to simplify our task.

The method described here for doing so follows closely the methods of
estimating photometric redshifts used by \citet{csabai00a} and
\citet{budavari00a}. It is designed to take advantage of what we
already understand about galaxy spectral energy distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.
Beginning with a space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), one can find an
orthonormal set of spectra spanning that space. To be more precise, we
define the dot-product in the space of galaxy spectra to be:
\begin{equation}
\vv{b}_i \cdot \vv{b}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
b_i(\lambda) b_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range corresponding to the range
which is within the observed wavelength range over most of the
redshift range of the sample. Otherwise, orthogonality is defined in
terms of regions of the spectra which are poorly constrained, which
make interpretation of the fits inconvenient.  This choice does not
affect the results for the reconstructed SED fit quantitatively
whatsoever.\footnote{If the template spectra were identical in the
subrange chosen and only differed outside of it, this choice would
cause computational problems, but in fact the spectra are all
independent in the restricted, as well as the full, wavelength range.}

We then further define the basis of the SED space such that the $N_b$
template SEDs, $\vv{v}_i$, are linear combinations of $N_b$ basis
spectra $\vv{b}_i$ and that $\vv{b}_i \cdot \vv{b}_j =
\delta_{ij}$. These conditions do not fully specify the $\vv{b}_i$,
which naturally may be rotated arbitrarily within our
$N_b$-dimensional subspace of spectral space.  In our implementation,
we use the subspace defined by linear combinations of ten
Bruzual-Charlot instantaneous burst models (our $\vv{v}_i(\lambda)$)
with ages ranging from $3 \times 10^7$ to $2\times 10^{10}$ years,
five with metallicity $Z=0.02$ and six with $Z=0.004$, all assuming a
Salpeter Initial Mass Function and smoothed using a Gaussian with a
standard deviation of $\sigma = 300$\AA. The smoothing is necessary
to prevent the method from overfitting; not all the small-scale
information about the templates can be recovered using this method,
and we find that this level of smoothing gives sensible results.

We cannot determine an individual galaxy's position in this
ten-dimensional space from its broad-band colors alone, because there
are only five observed bands. However, as outlined by
\citet{csabai00a}, one can find a lower dimensional space defined by
$N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
galaxies. The SED of galaxy $i$ may be approximated as a linear
combination of the basis spectra in this lower dimensional space:
\begin{equation}
\label{lincomb}
f_{\mathrm{model},i}(\lambda) d\lambda = 
\sum_k a_{i,k} \sum_j e_{k,j} b_j(\lambda) d\lambda,
\end{equation}
where $a_k$ are the components of the galaxy projected in the
low-dimensional space. From this SED it is easy to determine the flux
$F_{\mathrm{model},il}$ measured in each survey bandpass by projecting
the model SED onto the bandpass $l$. We define
\begin{equation}
\chi^2 = \sum_i \sum_l
\frac{(F_{\mathrm{obs},il}-F_{\mathrm{model},il})^2}{\sigma_{il}^2},
\end{equation}
where $\sigma_{il}$ is the estimated error in the observed flux
$F_{\mathrm{obs},il}$ in bandpass $l$ for galaxy $i$. Taking
derivatives of $\chi^2$ with respect to $a_{i,k}$ and $e_{k,j}$
results in a set of equations bilinear in $a_{i,k}$ and $e_{k,j}$.
\citet{csabai00a} describe how to iteratively solve this bilinear
equation by first fixing $e_{k,j}$ and fitting for $a_{i,k}$, then
fixing $a_{i,k}$ and fitting for $e_{k,j}$, and iterating that
procedure.

@@

We define the ``flux'' $f_k$ of each template as the flux in the range
$\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
estimate of the flux in this range is $F_{{v}}=\Sigma_k a_k f_k$. The
locus of points of a fixed flux in the component space $a_k$ lie on a
plane. Therefore, a natural choice for the orientation of the axes
$\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
constant flux; in this manner, the coefficient $a_0$ in Equation
(\ref{lincomb}) is directly proportional to $F$ and is thus purely a
measure of the object's flux (in the wavelength range considered),
while the parameters $a_i/a_0$ (where $i>0$) primarily measure the
shape of the galaxy SED in the wavelength range considered. $F$ can be
trivially related to $L$, the luminosity in this range, by the inverse
square of the luminosity distance.  As it turns out, the parameters
$a_i/a_0$ tend to be distributed very approximately in an ellipsoid,
with not much curvature within the space defined by the template
SEDs. Thus, we rotate the axes $\vv{b}_i$ (for $i>0$) such that they
are aligned with the principal axes of the ellipsoid. This choice of
coordinate system is arbitrary, however, and does not affect our
quantitative results.

@@

The version of the $K$-correction code (v1\_0) implementing the
calculations described here, along with eigentemplates and filter
curves is publicly available on the World Wide Web at {\tt
http://physics.nyu.edu/\~mb144/kcorrect}. In addition, a public CVS
repository can be accessed at {\tt xxx} (instructions for exporting
the appropriate version of the code from this repository are provided
on the web page noted above. The whole of the code can be used through
the Research Systems, Incorporated, IDL language; everything except
for the template-fitting also exists in a stand-alone C version (which
calls the same routines, guaranteeing consistency).

@@

The SDSS (\citealt{york00a}) is producing imaging and spectroscopic
surveys over $\pi$ steradians in the Northern Galactic Cap. A
dedicated 2.5m telescope (Siegmund {\it et al.}, in preparation) at
Apache Point Observatory, Sunspot, New Mexico, images the sky in five
bands between 3000 and 10000 \AA\ ($u$, $g$, $r$, $i$, $z$;
\citealt{fukugita96a}) using a drift-scanning, mosaic CCD camera
(\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
22.5$. The ultimate goal is to spectroscopically observe 900,000
galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Bright
Red Galaxies (BRGs; Eisenstein {\it et al.}, in preparation), and
100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation)
will be targeted for spectroscopic follow up using two digital
spectrographs on the same telescope. Many of the details of the survey
are described in a paper accompanying the Early Data Release
(\citealt{stoughton01a}). The survey has begun in earnest, and has
obtained about 20\% of its intended data.

@@

The magnitudes are calibrated $AB$, meaning that the system is
designed such that:
\begin{eqnarray}
m_{AB} &=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\cr
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons which are detected as a
function of wavelength (a unitless quantity). The difference in the
zeropoint is simply the factor of speed of light $c$ (expressed in
\AA\ s$^{-1}$) between the frequency and wavelength scales. {\bf
clearer on units}

The response functions $R(\lambda)$ have been measured using a
monochrometer by {\bf how ref Doi}.  Using a model for the atmospheric
transmission and the reflectivity of the primary and secondary
mirrors, one can then model the response of the entire system. For
each filter in the SDSS, there are six different CCDs; it has been
shown that the differences between these six camera columns are
negligible. The resulting set of filter curves is shown in Figure
\ref{response_sdss}, in comparison to a model of a galaxy spectral
energy distribution observed at $z=0$ (a 4 Gyr old instantaneous burst
from the models of \citealt{bruzual93a}).

@@

We use throughout the magnitudes determined by the SDSS using the
modified form of the magnitude described by \citet{petrosian76a}, as
described in \citet{strauss02a}, reddening-corrected using the dust
maps of \citet{schlegel98a}. These magnitudes have a nearly constant
metric aperture as a function redshift; in addition, all the bands use
the same aperture, so the measured SED corresponds (to within the
effects of seeing) to the SED of an identifiable region of the galaxy.

@@

For the purposes of applying the method described in Section
\ref{sedfit} to SDSS data, we take a subsample of the data consisting
of around 25,000 objects in the range $0.0<z<0.5$. The sample includes
both the main sample and the LRGs; we sample these spectra such that
we get an approximately even distribution within our range of
redshifts. In addition, we add results from galaxies in several plates
(totalling about 1,000 objects) which were selected by a photometric
redshift algorithm ({\bf ref who?}) to be at around $z\sim 0.3$. These
objects are invaluable for tying down the blue end of the templates.

@@

We choose to fit for $N_t = 4$ eigentemplates, the maximum one can use
and still fit allow freedom to fit for the templates themselves. Since
galaxies live in a low-dimensional space, we find it is worth being
able to fit for the templates. Simply using five templates (which
obviously reproduces all the magnitudes exactly) tends to yield
unphysical trends of galaxy SED versus redshift ({\it cf.}  Figure
\ref{constantz} below). Using three templates does nearly as well as
four templates in the sense that the resulting templates reproduce the
magnitudes nearly as well. However, since one of the applications of
these SED determinations is the distribution of galaxy colors in
fixed-frame magnitudes, we don't want to artificially reduce the
dimensionality of the color space to only two.

One more choice needs to be made, the wavelength regime over which to
orthogonalize the templates and over which to calculate the flux. We
choose the range defined by $\lambda_{\mathrm{min}}=3500$\AA and
$\lambda_{\mathrm{max}}=7500$\AA, since this range is included for
almost all galaxies in the sample. We refer here and in other papers
to the flux and luminosity in this range as the ``visual flux'' $f_v$
and the ``visual luminosity'' $L_v$.

@@

%The resulting four templates are shown in Figure \ref{k_espec_plot} in
%the range $2000<\lambda<12000$\AA, the region which is constrained by
%our observations. 
The top two panels of Figure \ref{k_coeffdist_plot} show the pairwise
joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$. In the
bottom panel, three spectra taken from a one-dimensional sequence
along the galay locus are shown, showing that the spectra become
progressively bluer along that sequence. Note that constraints on the
SED become poor at the bluest and reddest edges of this diagram (the
peak of the $z$-band response is at 8700\AA). Therefore, the odd
behavior near the edges (particularly at the red edge) should not be
taken too seriously.

These reconstructed spectra do an excellent job of reproducing the
observed galaxy fluxes. Figure \ref{model} shows the differences
between the observed and reconstructed fluxes as a function of
redshift and as a function of absolute magnitude (in $\band{0.1}{r}$).
The solid lines show the median differences. The estimates of the
fluxes in $g$, $r$, $i$, and $z$ do not correlate strongly with any of
these parameters, while for $u$ there is a 5--10\% discrepancy for the
most distant galaxies, in the sense that the model $u$-band flux is
too bright. The dotted lines show the boundaries containing 80\% of
the objects surrounding the median (roughly the $1.3\sigma$ limits if
the distribution were Gaussian). For $g$, $r$, and $i$, the scatter is
very small --- certainly comparable to the errors in calibration ---
and roughly independent of redshift and absolute magnitude. For $z$,
the scatter is somewhat larger, perhaps reflecting the poor
signal-to-noise in $z$, but remains constant with redshift. For $u$,
the scatter is larger, perhaps reflecting the fact that four templates
simply is not enough to characterize galaxy SEDs simultaneously in the
optical and near ultra-violet. After all, galaxies are known to have
large variations in their near-ultraviolet properties (because it
depends strongly on dust content and recent star-formation). This
effect may become more important with redshift as $u$ probes bluer and
bluer regions of the SED. All told, these spectra seem to be fair
estimates of the true underlying spectra.

An important test of the reasonableness of the fits is to check that
for a fixed absolute magnitude, the distribution of the SED fits
depends little on redshift. Figures \ref{coeffs1}--\ref{coeffs4}
demonstrates this to be the case for galaxies in several luminosity
ranges. The left column of Figure \ref{coeffs1} shows the three
coefficient ratios $a_i/a_0$ for each galaxy as a function of
redshift. The right column shows the distribution of these coefficient
ratios for two redshift ranges, showing that they are nearly
consistent. The resulting change in the color distribution inferred
from these coefficients is similarly small, at the level of what one
would expect intrinsic evolution to cause.

@@

With a model of the rest-frame galaxy SED in hand, it is 
straightforward to project that SED onto a filter corresponding to 
an observation at any redshift, and predict the so-called
$K$-correction to the galaxy magnitude in order to compare galaxies 
at different redshifts. One simply calculates the AB magnitude at 
any redshift as:
\begin{equation}
m_{AB} = -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}
f(\frac{\lambda}{1+z}) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right].
\end{equation}
The simple way to think about this equation is that simply moving away
from the Earth doing does not change the bolometric flux of the
galaxy. So when the spectrum is stretched by the $(1+z)$ redshift
factor, as in $f(\lambda/(1+z))$, the bolometric flux should be
constant; that is, you must apply the extra multiplicative factor
$1/(1+z)$ to counteract the stretching of the spectrum. Once you have
have accounted for the fact that you are observing a particular
bandpass given the redshift of the object, you can then go ahead and
apply to the calculated magnitude the cosmological distance modulus
(that is, the luminosity distance-squared law, as tabulated by, {\it
e.g.} \citealt{hogg97a}).

>>>>

We implemented this system and are publishing it in order to prepare
the reader for upcoming papers which will rely heavily on the
reliability of the SED fits described here.  However, having
implemented the fits, there is no reason not to share them with the
community; thus, this paper serves the secondary purpose of describing
a piece of software being released to the astronomical community. The
computer software created is a C shared object library, around which
we have written both stand-alone C programs and IDL routines. We have
made the source code available publicly, through a URL on the World
Wide Web and through a public CVS repository. Improvements or ports to
other languages implemented by users may be incorporated into the code
upon request. The conditions of use for the code are that this paper
is cited in any resulting refereed journal article and that the
version of the code used is specified in any such paper. The version
of the code used to make the figures for this paper is {\tt kcorrect
v1\_2}.

@@

Our task is to recover a model for the galaxy SED from broad-band
photometric measurements. Since a set of broad band galaxy fluxes does
not correspond uniquely to a particular SED, and because galaxy SEDs
are known to have significant structure over wavelength ranges small
compared to our bandpass, this task is an ill-posed, inverse
problem. However, we are not completely ignorant about the forms which
galaxy SEDs take, so we can attempt to use what we know about galaxy
SEDs to simplify our task.  The method described here for doing so
follows closely the methods of estimating photometric redshifts used
by \citet{csabai00a} and \citet{budavari00a}. It is designed to take
advantage of what we already understand about galaxy spectral energy
distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.

Let us begin with an SED space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), where $N_b$ is
large. In principle, this should be a complete set of galaxy SEDs if
we want to express {\it any} galaxy SED in this basis set; in
practice, this property is not necessary.  Given a set of
$\vv{v}_i(\lambda)$, one can find an orthonormal set of spectra
$\vv{b}_i(\lambda)$ spanning the space.

To be more precise, we define the dot-product in the space of galaxy
spectra to be:
\begin{equation}
\vv{v}_i \cdot \vv{v}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
v_i(\lambda) v_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range corresponding to the range
which is within the observed wavelength range over most of the
redshift range of the sample. Otherwise, orthogonality is defined in
terms of regions of the spectra which are poorly constrained, which
make interpretation of the fits inconvenient.  This choice does not
affect the results for the reconstructed SED fit quantitatively
whatsoever.\footnote{If the template spectra were identical in the
subrange chosen and only differed outside of it, this choice would
cause computational problems, but in fact the spectra are all
independent in the restricted, as well as the full, wavelength range.}
We then define the basis of the SED space such that the $N_b$ template
SEDs, $\vv{v}_i$, are linear combinations of $N_b$ basis spectra
$\vv{b}_i$ and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These
conditions do not fully specify the $\vv{b}_i$, which naturally may be
rotated arbitrarily within our $N_b$-dimensional subspace of spectral
space.

In our case the number of observed bands is less than $N_b$, so we
cannot determine an individual galaxy's position in this
$N_b$-dimensional space from its broad-band colors alone.  However, as
outlined by \citet{csabai00a}, one can find a lower dimensional space
defined by $N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
galaxies, as follows. The SED of galaxy $i$ may be reconstructed from a
linear combination of the basis spectra in this lower dimensional
space:
\begin{equation}
\label{lincomb}
f_{\mathrm{rec},i}(\lambda) d\lambda = 
\sum_k a_{i,k} \sum_j e_{k,j} b_j(\lambda) d\lambda,
\end{equation}
where $a_k$ are the components of the galaxy projected in the
low-dimensional space. From this SED it is easy to determine the
reconstructed flux
$F_{\mathrm{rec},il}$ in each survey bandpass by projecting
the model SED onto the bandpass $l$: 
\begin{equation}
F_{\mathrm{rec},il} = \int_0^\infty d\lambda f(\lambda) S(\lambda),
\end{equation}
where $S(\lambda)$ is the response of the instrument.  We define
\begin{equation}
\chi^2 = \sum_i \sum_l
\frac{(F_{\mathrm{obs},il}-F_{\mathrm{rec},il})^2}{\sigma_{il}^2},
\end{equation}
where $\sigma_{il}$ is the estimated error in the observed flux
$F_{\mathrm{obs},il}$ in bandpass $l$ for galaxy $i$. Taking
derivatives of $\chi^2$ with respect to $a_{i,k}$ and $e_{k,j}$
results in a set of equations bilinear in $a_{i,k}$ and $e_{k,j}$.
\citet{csabai00a} describe how to iteratively solve this bilinear
equation by first fixing $e_{k,j}$ and fitting for $a_{i,k}$, then
fixing $a_{i,k}$ and fitting for $e_{k,j}$, and iterating that
procedure.

@@

We define the ``total flux'' $f_k$ of each template as the flux in the
range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$. The
locus of points of a fixed flux in the component space $a_k$ lie on a
plane. Therefore, a natural choice for the orientation of the axes
$\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
constant flux; in this manner, the coefficient $a_0$ in Equation
(\ref{lincomb}) is directly proportional to $F$ and is thus purely a
measure of the object's flux (in the wavelength range considered),
while the parameters $a_i/a_0$ (where $i>0$) primarily measure the
shape of the galaxy SED in the wavelength range considered. $F$ can be
trivially related to $L$, the luminosity in this range, by the inverse
square of the luminosity distance.  As it turns out, the parameters
$a_i/a_0$ tend to be distributed very approximately in an ellipsoid,
with not much curvature within the space defined by the template
SEDs. Thus, we rotate the axes $\vv{e}_i$ (for $i>0$) such that they
are aligned with the principal axes of the ellipsoid. Furthermore, we
normalize the vectors $\vv{e}_i$ such that the $a_0=F_t$ in units of
ergs cm$^{-2}$ s$^{-1}$. This choice of coordinate system is
arbitrary and does not affect our quantitative results, but it is
convenient.

@@

The version of the $K$-correction code (v1\_2) implementing the
calculations described here, along with eigentemplates and filter
curves is publicly available on the World Wide Web at {\tt
http://physics.nyu.edu/\~\ mb144/kcorrect}. In addition, a public CVS
repository can be accessed at {\tt xxx} (instructions for exporting
the appropriate version of the code from this repository are provided
on the web page noted above. The whole of the code can be used through
the Research Systems, Incorporated, IDL language; everything except
for the template-fitting also exists in stand-alone C programs (which
calls the same routines, guaranteeing consistency).

@@

The SDSS (\citealt{york00a}) is producing imaging and spectroscopic
surveys over $\pi$ steradians in the Northern Galactic Cap. A
dedicated 2.5m telescope (Siegmund {\it et al.}, in preparation) at
Apache Point Observatory, Sunspot, New Mexico, images the sky in five
bands between 3000 and 10000 \AA\ ($u$, $g$, $r$, $i$, $z$;
\citealt{fukugita96a}) using a drift-scanning, mosaic CCD camera
(\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
22.5$. The ultimate goal is to spectroscopically observe 900,000
galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Bright
Red Galaxies (BRGs; Eisenstein {\it et al.}, in preparation), and
100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation)
will be targeted for spectroscopic follow up using two digital
spectrographs on the same telescope. Many of the details of the galaxy
survey are described in a description of the galaxy target selection
paper (\citealt{strauss02a}). Other aspects of the survey are
described in the Early Data Release (EDR; \citealt{stoughton01a}). The
survey has begun in earnest, and has obtained about 20\% of its
intended data.

@@

The magnitudes are calibrated $AB$, meaning that the system is
designed such that:
\begin{eqnarray}
m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\cr
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons entering the Earth's
atmosphere which are detected as a function of wavelength (a unitless
quantity). In this equation, $f(\lambda)$ is in units of ergs
cm$^{-2}$ s$^{-1}$ \AA$^{-1}$, and $f(\nu)$ is in units of ergs
cm$^{-2}$ s$^{-1}$ Hz$^{-1}$. The normalizations defined here mean
that an object with $f(\nu) = 3.631 \times 10^{-20}$ ergs cm$^{-2}$
s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes equal to zero.  The
difference in the zeropoint in the two equations is simply the factor
of speed of light $c$ (expressed in \AA\ s$^{-1}$) between the
frequency and wavelength scales.

The response functions $R(\lambda)$ have been measured using a
monochrometer by {\bf how ref Doi}.  Using a model for the atmospheric
transmission and the reflectivity of the primary and secondary
mirrors, one can then model the response of the entire system. For
each filter in the SDSS, there are six different CCDs; it has been
shown that the differences between these six camera columns are
small. The resulting set of filter curves is shown in Figure
\ref{response_sdss}, in comparison to a model of a galaxy spectral
energy distribution observed at $z=0$ (a 4 Gyr old instantaneous burst
from the models of \citealt{bruzual93a}).

@@

We use two types of magnitudes determined by the SDSS. First, the
modified form of the magnitude described by \citet{petrosian76a}, as
described in \citet{strauss02a}.  These magnitudes have a nearly
constant metric aperture as a function redshift; in addition, all the
bands use the same aperture, so the measured SED corresponds (to
within the effects of seeing) to the SED of an identifiable region of
the galaxy. However, for faint objects, the Petrosian magnitudes tend
to become noisy. Thus, for galaxies with $z>0.25$ we instead use the
higher signal-to-noise ``model magnitudes.'' Model magnitudes are
calculated in all bands using a single weighted aperture. The weighted
aperture is the better fitting model (pure exponential or pure de
Vaucouleurs) to the galaxy image in the $r$-band. In
\citet{stoughton01a} we are explicitly warned (in paragraphs written
by the lead author of this paper!)  not to mix Petrosian and model
magnitudes in a single analysis, since they measure galaxies in very
different ways. Nevertheless, for the purposes of contraining galaxy
template SEDs, it is perfectly acceptable to use {\it any}
well-defined part of any galaxy.  We reddening-correct both types of
magnitudes using the dust maps of \citet{schlegel98a}.

@@

For the purposes of applying the method described in Section
\ref{sedfit} to SDSS data, we take a subsample of the data consisting
of around 25,000 objects in the range $0.0<z<0.5$. The sample includes
both the main sample and the LRGs; we sample these spectra such that
we get an approximately even distribution within our range of
redshifts. In addition, we add results from galaxies in several plates
(totalling about 1,000 objects) which were selected by a photometric
redshift algorithm ({\bf ref who?}) to be at around $z\sim
0.3$--$0.4$. These objects are invaluable for tying down the blue end
of the templates.

@@

For the SED space (our $\vv{v}_i(\lambda)$), we use the subspace
defined by linear combinations of ten Bruzual-Charlot instantaneous
burst models with ages ranging from $3 \times 10^7$ to $2\times
10^{10}$ years, five with metallicity $Z=0.02$ and five with
$Z=0.004$, all assuming a Salpeter Initial Mass Function. We find that
we cannot recover information about the SED below a certain scale,
except perhaps around the 4000\AA\ break. Thus, we smooth most of the
wavelength regime of the templates using a Gaussian with a standard
deviation of $\sigma = 300$\AA; around 4000\AA, we smooth only with
$\sigma = 150$\AA.

We choose to fit for $N_t = 4$ eigentemplates, the maximum one can use
and still fit allow freedom to fit for the templates themselves. Since
galaxies live in a low-dimensional space, we find it is worth being
able to fit for the templates. Simply using five templates (which
obviously reproduces all the magnitudes exactly) tends to yield
unphysical trends of galaxy SED versus redshift ({\it cf.}  Figure
\ref{constantz} below). Using three templates does nearly as well as
four templates in the sense that the resulting templates reproduce the
$griz$ magnitudes nearly as well. However, the fourth template is
necessary to recover the $u$-band flux to better than about 15\%. In
addition, since one of the applications of these SED determinations is
the distribution of galaxy colors in fixed-frame magnitudes, we don't
want to artificially reduce the dimensionality of the color space to
only two.

One more choice needs to be made, the wavelength regime over which to
orthogonalize the templates and over which to calculate the flux. We
choose the range defined by $\lambda_{\mathrm{min}}=3500$\AA\ and
$\lambda_{\mathrm{max}}=7500$\AA, since this range is included for
almost all galaxies in the sample. We refer here and in other papers
to the flux and luminosity in this range as the ``visual flux'' $f_v$
and the ``visual luminosity'' $L_v$.

@@

%The resulting four templates are shown in Figure \ref{k_espec_plot} in
%the range $2000<\lambda<12000$\AA, the region which is constrained by
%our observations. 
The top two panels of Figure \ref{k_coeffdist_plot} show the pairwise
joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$. In the
bottom panel, three spectra taken from a one-dimensional sequence
along the galay locus are shown, showing that the spectra become
progressively bluer along that sequence. Note that constraints on the
SED become poor at the bluest and reddest edges of this diagram (the
peak of the $z$-band response is at 8700\AA). Therefore, the odd
behavior near the edges should not be taken too seriously.

These reconstructed spectra do an excellent job of reproducing the
observed galaxy fluxes. Figure \ref{k_model_plot} shows the
differences between the observed and reconstructed fluxes as a
function of redshift and as a function of absolute magnitude (in
$\band{0.1}{r}$).  The solid lines show the median differences. The
estimates of the fluxes in $g$, $r$, $i$, and $z$ do not correlate
strongly with any of these parameters, while for $u$ there is a
5--10\% discrepancy for the most distant galaxies, in the sense that
the model $u$-band flux is too bright. The dotted lines show the
boundaries containing 80\% of the objects surrounding the median
(roughly the $1.3\sigma$ limits if the distribution were
Gaussian). For $g$, $r$, and $i$, the scatter is very small ---
certainly comparable to the errors in calibration --- and roughly
independent of redshift and absolute magnitude. For $z$, the scatter
is somewhat larger, perhaps reflecting the poor signal-to-noise in
$z$, but remains constant with redshift. For $u$, the scatter is
larger, perhaps reflecting the fact that four templates simply is not
enough to characterize galaxy SEDs simultaneously in the optical and
near ultra-violet. After all, galaxies are known to have large
variations in their near-ultraviolet properties (because it depends
strongly on dust content and recent star-formation). This effect may
become more important with redshift as $u$ probes bluer and bluer
regions of the SED. All told, these spectra seem to be fair estimates
of the true underlying spectra.

An important test of the reasonableness of the fits is to check that
for a fixed type of galaxy, the distribution of the SED fits
depends little on redshift. We must be careful in performing this
check that we are comparing Figures \ref{coeffs1}--\ref{coeffs4}
demonstrates this to be the case for galaxies in several luminosity
ranges. The left column of Figure \ref{coeffs1} shows the three
coefficient ratios $a_i/a_0$ for each galaxy as a function of
redshift. The right column shows the distribution of these coefficient
ratios for two redshift ranges, showing that they are nearly
consistent. The resulting change in the color distribution inferred
from these coefficients is similarly small, at the level of what one
would expect intrinsic evolution to cause.

@@

With a model of the rest-frame galaxy SED in hand, it is 
straightforward to project that SED onto a filter corresponding to 
an observation at any redshift, and predict the so-called
$K$-correction to the galaxy magnitude in order to compare galaxies 
at different redshifts. One simply calculates the AB magnitude at 
any redshift as:
\begin{equation}
m_{AB} = -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}
f(\frac{\lambda}{1+z}) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right].
\end{equation}
The simple way to think about this equation is that simply moving away
from the Earth doing does not change the bolometric flux of the
galaxy. So when the spectrum is stretched by the $(1+z)$ redshift
factor, as in $f(\lambda/(1+z))$, the bolometric flux should be
constant; that is, you must apply the extra multiplicative factor
$1/(1+z)$ to counteract the stretching of the spectrum. Once you have
have accounted for the fact that you are observing a particular
bandpass given the redshift of the object, you can then go ahead and
apply to the calculated magnitude the cosmological distance modulus
(that is, the luminosity distance-squared law, as tabulated by, {\it
e.g.} \citealt{hogg97a}).
